Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-01-15 vesicle_code/docs/adhesion/adhesion.tex
fixed some latex fomratting issues.

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[aps,prb,showpacs,amsmath,amssymb,superscriptaddress]{revtex4-1}

@@

\usepackage[titletoc]{appendix}
\usepackage{todonotes}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{stmaryrd}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\author{Bryan Quaife}
\affiliation{Department of Scientific Computing, Florida State University, Tallahassee, Florida, USA}
% \email{bquaife@fsu.edu}
% \altaffiliation[]{Department of Scientific Computing \\ Florida State University}
\author{Shravan Veerapaneni}%
% \email{shravan@umich.edu}
\affiliation{Department of Mathematics, University of Michigan, Ann Arbor, Michigan, USA}%
\author{Y.-N. Young}%
 \email[Corresponding author: ]{yyoung@njit.edu}
\affiliation{Department of Mathematical Sciences, Newark, NJ, USA}%

%\collaboration{MUSO Collaboration}%\noaffiliation

%\author{Charlie Author}
% \homepage{http://www.Second.institution.edu/~Charlie.Author}
%\affiliation{
% Second institution and/or address\\
% This line break forced% with \\
%}%
%\affiliation{
% Third institution, the second for Charlie Author
%}%
%\author{Delta Author}
%\affiliation{%
% Authors' institution and/or address\\
% This line break forced with \textbackslash\textbackslash
%}%
%
%\collaboration{CLEO Collaboration}%\noaffiliation

@@

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been widely utilized
as biological cell mimics in biophysics and material engineering \cite{}. 
%
The ever expanding applications have encouraged detailed experimental,
theoretical and numerical investigations of vesicle hydrodynamics in
applied flow, electric and magnetic fields. Theoretical investigations
often take the form of small-deformation analysis or spheroidal dynamics
of a nearly-spherical vesicle (or nearly-circular in two dimensions)
subjected to an applied flow or under an external
forcing~\cite{Barthes-BieselRallison1981_JFM,Misbah2006_PRL,Vlahovska2007_PRE,Finken2008_EPL,ZhangZahnTanLin2013_PoF,Nganguia2013_PRE}.
On the other hand, various numerical methods have been developed for
simulating the transient hydrodynamics of vesicle
suspensions~\cite{BagchiJohoson2005_JBE,Biben2005_EJP,Veerapaneni2009_JCP,SeolHuKimLai2016_JCP}.  
%Experiments are performed to study vesicle hydrodynamics under various flowing conditions or forcing configurations \cite{MaderVitkova2006_EurPhysJE, Abkarian2008_SoftMatt,DahlNarsimhanGouveia2016_SoftMatt}.

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by excess area (reduced volume), viscosity contrast
between interior and exterior fluids, and shear rate of the imposed
far-field fluid flow. In addition a vesicle with a rigid particle inside
is also investigated as a biological mimic of a eukaryotic cell with a
nucleus that occupies nearly half of the intracellular
volume~\cite{Veerapaneni2011_PRL}. Small-deformation analysis shows that
a vesicle tank-treads in a planar shear flow for low viscosity contrast
and shear rate. At high viscosity contrast the tank-tread (TT) dynamics
transitions to tumbling (TB)~\cite{Misbah2006_PRL,Vlahovska2007_PRE},
and this leads to a transition in effective shear viscosity in the
vesicle suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also
validated by direct numerical
simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,ZabuskySegreDeschamps2011_PoF}.
Between TT and TB vesicle hydrodynamics, a breathing (tremble) is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,ZhaoShaqfeh2011_JFM,SpannZhaoShaqfeh2014_PoF}
to affect the effective shear viscosity in a different way.  In an
extensional flow (planar or uniaxial), vesicle shape dynamics depends
sensitively on the vesicle reduced volume  and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL,ZhaoShaqfeh2013_JFM,Narsimhan2014_JFM,DahlNarsimhanGouveia2016_SoftMatt}:
Asymmetric shape and oscillatory undulation of the vesicle membrane are
two examples of the complex vesicle hydrodynamics in an extensional
flow.

Collective dynamics of vesicles naturally is dictated by the individual vesicle-vesicle interactions.
%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
In numerical simulations of vesicle suspensions, it is often assumed
that the liquid film between vesicles acts to keep the membrane-membrane
interaction, such as the van der Waals type adhesive interaction,
negligible.  However the amphiphilic natures of lipid molecules (a
hydrophobic tail and a hydrophilic head with an electric dipole)
complicate the interactions between a lipid bilayer membrane and another
bilayer
membrane~\cite{EvansMetcalfe1984_BJ,Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces}
or a solid (such as a glass substrate or a nano particle).  Studies
confirmed that such interaction potential inevitably consists of a
long-range attraction component and a short-range repulsion
component~\cite{Book_IntermolecularSurfaceForces}, and the combination
of the two gives rise to adhesion at the separation distance that
minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. Such adhesive
potentials are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is essential for membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.

%
Adhesion between a vesicle and a substrate has been extensively studied,
primarily focused on analyzing the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase}
than the transient adhesion process; exceptions
include~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ}, who measured the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker
constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds
to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion
energy per unit area to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes under adhesion in these regimes
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

This paper is organized as follows. In \S~\ref{sec:ge} we formulate our
model for vesicle hydrodynamics with adhesive interactions  between
membranes of distinct vesicles.  We simulate the adhesion process of two
vesicles in a quiescent flow in \S~\ref{sec:qflow}, where we also
present a simple model to estimate how long it takes for two nearby
vesicles to reach an equilibrium configuration under adhesion. In
\S~\ref{sec:eflow} we numerically investigate the hydrodynamics of a
vesicle doublet in a fluid trap where the stagnation point is actively
controlled to be at the center of the vesicle doublet. From these
results we propose a novel application of the microfluidic fluid trap to
probe the adhesion strength between lipid bilayer membranes in
solutions. In \S~\ref{sec:sflow} we investigate how two vesicles may
form a doublet as they move toward each other under a planar shear flow.
We map out the parameter regions for bound/unbound vesicles under a
planar shear flow, and we also investigate the effects of adhesive
interactions on the rheological properties of a dilute suspension of
vesicle doublets.  Finally in \S~\ref{sec:conclusions} we discuss the
implications of our results and point out a few potential future
directions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, an integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2,
  \label{eqn:SLP}
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$.  The traction jump of the
velocity $\uu(\xx) = \SS[\ff](\xx)$ is the density function $\ff$, so
the fluid velocity is $\xx \in \RR^2$ is
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx).
\end{align*}
The traction jump is given by three terms: the bending, tension, and
adhesion.  Therefore
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx.
\end{align*}
Applying the no-slip boundary condition on each vesicle, the governing
equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition can be written as
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{qua-bir2014}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
  \label{eqn:IMEXEuler}
\end{align}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply
Alpert quadrature to the logarithmically singular single-layer potential
$\SS$~\cite{alp1999}.  The source and target points of the adhesion
force never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.  To control the error and
achieve higher-order accuracy in time, we apply a time adaptive spectral
deferred correction method~\cite{qua-bir2016}.  The SDC method is able
to achieve second-order accuracy by iteratively applying the first-order
method~\eqref{eqn:IMEXEuler}.

@@

When two vesicles move towards (or away from) each other under a
constant force $F$ without any external fluid flow, the height $(h)$ of
the thin liquid film between two vesicles follows the draining
dynamics~\cite{RamachandranLeal2010_PoF}
\begin{equation}
\label{eq:dhdt}
\frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{equation}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of
exterior fluid, $R_0$ is the radius of the undeformed vesicle.  For the
case of a constant forcing $F$ (independent of separation distance $h$
and time $t$), equation~\eqref{eq:dhdt} can be easily integrated to give
a relation between film thickness $h$ and time $t$:
\begin{equation}
t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}\frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{equation}
where $h(0)$ is the vesicle separation at $t=0$.
We note that $F< (>) 0$ for an attractive  (repulsive) interaction
between two vesicles,  and consequently $h(t)$ decreases (increases)
from  the initial separation distance $h(0)$.  When the force is
attractive, $h$ decreases monotonically and $t$ in the above equation is
the draining time that diverges as $h\rightarrow 0$.
%Similarly $F>0$ for a repulsive force and consequently $h(t) > h(0)$.

When the force on each vesicle is a function of separation distance $h$ of the form (with an integer $m\ge 2$)
\begin{equation}
\label{eq:Fh}
F(h) \sim H \frac{\delta^m}{h^{m+1}}\left(\left(\frac{\delta}{h}\right)^2-1\right),
\end{equation}
attraction is dominant at ``large" scales ($h > \delta$) while repulsion
is dominant at ``small" scales ($h<\delta$).  Integration of
equation~\eqref{eq:dhdt} with the above $F(h)$ gives the relationship
between $t$ and $h$.  In this case the relationship involves an integral
of a function of dimensionless $\bar{h} \equiv h/\delta$:
\begin{align}
t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = \int^{h(t)}_{h(0)}-\frac{dh}{h^3
  F^{1/3}} = -\frac{1}{H^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}\frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^2-1)^{1/3}},
\end{align}
\begin{equation}
\label{eq:teq_scaling}
t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{H^{1/3}\delta^{5/3}}.
\end{equation}
Equation~\eqref{eq:teq_scaling} tells us that both adhesion strength $H$
and the separation distance $\delta$ affect the time it takes for a pair
of vesicles to reach equilibrium under adhesion force $F$ in
equation~\eqref{eq:Fh}.  In particular, the equilibration time $t$ is
proportional to (1) the separation distance $\delta^{-5/3}$ (thus
$t\rightarrow \infty$ as $\delta\rightarrow 0$), and (2) the adhesion
strength $H^{-1/3}$ (as in constant forcing
case~\cite{RamachandranLeal2010_PoF}).  From the above analysis we also
expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength
is independent of the specific form of the potential.

@@

To test the scaling of the equilibration time $t$ with respect to $H$
and $\delta$, we simulate the vesicle adhesion dynamics in a quiescent
flow.  Starting with two vesicles at a distance of twice the radius at
$t=0$, the long-range attraction pulls the vesicles together.
Figure~\ref{fig:qflow00}(a) plots the scaled adhesion energy versus time
with $H=0.1$ and three values of $\delta$ as labeled.  Adhesion energy
reaches minimum at equilibrium, and the scaled $E_{adh}$ evolves towards
zero at equilibrium.  The arrow indicate the times when the scaled
adhesion energy reaches within $1\%$ of equilibrium: $t\sim 96$ for
$\delta=0.2$, $t\sim 45$ for $\delta = 0.3$, and $t\sim 30$ for
$\delta=0.4$:
\begin{align}
\frac{t_{\delta=0.2}}{t_{\delta=0.4}} \sim \frac{96}{30}=3.2 &\Longleftrightarrow \left(\frac{0.4}{0.2}\right)^{5/3}\sim 3.17,\\
\frac{t_{\delta=0.2}}{t_{\delta=0.3}} \sim \frac{96}{45}=2.1 &\Longleftrightarrow \left(\frac{0.3}{0.2}\right)^{5/3}\sim 2.
\end{align}

>>>>

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
%\documentclass[aps,prb,showpacs,amsmath,amssymb,superscriptaddress]{revtex4-1}
\documentclass[prf,superscriptaddress,showpacs]{revtex4-1}

@@

\usepackage{amsmath,amssymb}
\usepackage[titletoc]{appendix}
\usepackage{todonotes}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{stmaryrd}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\author{Bryan Quaife}
\affiliation{Department of Scientific Computing, Florida State University, Tallahassee, Florida, USA}
% \email{bquaife@fsu.edu}
\author{Shravan Veerapaneni}%
% \email{shravan@umich.edu}
\affiliation{Department of Mathematics, University of Michigan, Ann Arbor, Michigan, USA}%
\author{Y.-N.~Young}%
 \email[Corresponding author: ]{yyoung@njit.edu}
\affiliation{Department of Mathematical Sciences, Newark, NJ, USA}%

@@

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been
widely utilized as biological cell mimics in biophysics and material
engineering \cite{}.  The ever expanding applications have encouraged
detailed experimental, theoretical and numerical investigations of
vesicle hydrodynamics in applied flow, electric and magnetic fields.
Theoretical investigations often take the form of small-deformation
analysis or spheroidal dynamics of a nearly-spherical vesicle (or
nearly-circular in two dimensions) subjected to an applied flow or under
an external forcing~\cite{Barthes-BieselRallison1981_JFM,
Misbah2006_PRL, Vlahovska2007_PRE, Finken2008_EPL,
ZhangZahnTanLin2013_PoF, Nganguia2013_PRE}.  On the other hand, various
numerical methods have been developed for simulating the transient
hydrodynamics of vesicle suspensions~\cite{BagchiJohoson2005_JBE,
Biben2005_EJP, Veerapaneni2009_JCP, SeolHuKimLai2016_JCP}.  
%Experiments are performed to study vesicle hydrodynamics under various flowing conditions or forcing configurations \cite{MaderVitkova2006_EurPhysJE, Abkarian2008_SoftMatt,DahlNarsimhanGouveia2016_SoftMatt}.

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by excess area (reduced volume), viscosity contrast
between interior and exterior fluids, and shear rate of the imposed
far-field fluid flow. In addition a vesicle with a rigid particle inside
is also investigated as a biological mimic of a eukaryotic cell with a
nucleus that occupies nearly half of the intracellular
volume~\cite{Veerapaneni2011_PRL}. Small-deformation analysis shows that
a vesicle tank-treads in a planar shear flow for low viscosity contrast
and shear rate. At high viscosity contrast the tank-tread (TT) dynamics
transitions to tumbling (TB)~\cite{Misbah2006_PRL, Vlahovska2007_PRE},
and this leads to a transition in effective shear viscosity in the
vesicle suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also
validated by direct numerical
simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,
ZabuskySegreDeschamps2011_PoF}.  Between TT and TB vesicle
hydrodynamics, a breathing (tremble) is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume  and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
Collective dynamics of vesicles naturally is dictated by the individual
vesicle-vesicle interactions.  In numerical simulations of vesicle
suspensions, it is often assumed that the liquid film between vesicles
acts to keep the membrane-membrane interaction, such as the van der
Waals type adhesive interaction, negligible.  However the amphiphilic
natures of lipid molecules (a hydrophobic tail and a hydrophilic head
with an electric dipole) complicate the interactions between a lipid
bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,
Book_IntermolecularSurfaceForces} or a solid (such as a glass substrate
or a nano particle).  Studies confirmed that such interaction potential
inevitably consists of a long-range attraction component and a
short-range repulsion component~\cite{Book_IntermolecularSurfaceForces},
and the combination of the two gives rise to adhesion at the separation
distance that minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. Such adhesive
potentials are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is essential for membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.

Adhesion between a vesicle and a substrate has been extensively studied,
primarily focused on analyzing the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase}
than the transient adhesion process; exceptions
include~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ}, who measured the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker
constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds
to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion
energy per unit area to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes under adhesion in these regimes
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

This paper is organized as follows. In Section~\ref{sec:ge} we formulate
our model for vesicle hydrodynamics with adhesive interactions  between
membranes of distinct vesicles.  We simulate the adhesion process of two
vesicles in a quiescent flow in Section~\ref{sec:qflow}, where we also
present a simple model to estimate how long it takes for two nearby
vesicles to reach an equilibrium configuration under adhesion. In
Section~\ref{sec:eflow} we numerically investigate the hydrodynamics of
a vesicle doublet in a fluid trap where the stagnation point is actively
controlled to be at the center of the vesicle doublet. From these
results we propose a novel application of the microfluidic fluid trap to
probe the adhesion strength between lipid bilayer membranes in
solutions. In Section~\ref{sec:sflow} we investigate how two vesicles
may form a doublet as they move toward each other under a planar shear
flow.  We map out the parameter regions for bound/unbound vesicles under
a planar shear flow, and we also investigate the effects of adhesive
interactions on the rheological properties of a dilute suspension of
vesicle doublets.  Finally in Section~\ref{sec:conclusions} we discuss
the implications of our results and point out a few potential future
directions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, an integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2,
  \label{eqn:SLP}
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$.  The traction jump of the
velocity $\uu(\xx) = \SS[\ff](\xx)$ is the density function $\ff$, so
the fluid velocity is $\xx \in \RR^2$ is
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx).
\end{align*}
The traction jump is given by three terms: the bending, tension, and
adhesion.  Therefore
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx.
\end{align*}
Applying the no-slip boundary condition on each vesicle, the governing
equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition can be written as
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{quaife2016adaptive}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
  \label{eqn:IMEXEuler}
\end{align}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply
Alpert quadrature to the logarithmically singular single-layer potential
$\SS$~\cite{alp1999}.  The source and target points of the adhesion
force never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.  To control the error and
achieve higher-order accuracy in time, we apply a time adaptive spectral
deferred correction method~\cite{quaife2016adaptive}.  The SDC method is
able to achieve second-order accuracy by iteratively applying the
first-order method~\eqref{eqn:IMEXEuler}.

@@

When two vesicles move towards (or away from) each other under a
constant force $F$ without any external fluid flow, the height $(h)$ of
the thin liquid film between two vesicles follows the draining
dynamics~\cite{RamachandranLeal2010_PoF}
\begin{align}
\label{eq:dhdt}
\frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{align}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of
exterior fluid, $R_0$ is the radius of the undeformed vesicle.  For the
case of a constant forcing $F$ (independent of separation distance $h$
and time $t$), equation~\eqref{eq:dhdt} can be easily integrated to give
a relation between film thickness $h$ and time $t$:
\begin{align*}
t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}\frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{align*}
where $h(0)$ is the vesicle separation at $t=0$.
We note that $F< (>) 0$ for an attractive  (repulsive) interaction
between two vesicles,  and consequently $h(t)$ decreases (increases)
from  the initial separation distance $h(0)$.  When the force is
attractive, $h$ decreases monotonically and $t$ in the above equation is
the draining time that diverges as $h\rightarrow 0$.
%Similarly $F>0$ for a repulsive force and consequently $h(t) > h(0)$.

When the force on each vesicle is a function of separation distance $h$ of the form (with an integer $m\ge 2$)
\begin{align}
\label{eq:Fh}
F(h) \sim H \frac{\delta^m}{h^{m+1}}\left(\left(\frac{\delta}{h}\right)^2-1\right),
\end{align}
attraction is dominant at ``large" scales ($h > \delta$) while repulsion
is dominant at ``small" scales ($h<\delta$).  Integration of
equation~\eqref{eq:dhdt} with the above $F(h)$ gives the relationship
between $t$ and $h$.  In this case the relationship involves an integral
of a function of dimensionless $\bar{h} \equiv h/\delta$:
\begin{align*}
t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = \int^{h(t)}_{h(0)}-\frac{dh}{h^3
  F^{1/3}} = -\frac{1}{H^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}\frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^2-1)^{1/3}},
\end{align*}
\begin{align}
\label{eq:teq_scaling}
t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{H^{1/3}\delta^{5/3}}.
\end{align}
Equation~\eqref{eq:teq_scaling} tells us that both adhesion strength $H$
and the separation distance $\delta$ affect the time it takes for a pair
of vesicles to reach equilibrium under adhesion force $F$ in
equation~\eqref{eq:Fh}.  In particular, the equilibration time $t$ is
proportional to (1) the separation distance $\delta^{-5/3}$ (thus
$t\rightarrow \infty$ as $\delta\rightarrow 0$), and (2) the adhesion
strength $H^{-1/3}$ (as in constant forcing
case~\cite{RamachandranLeal2010_PoF}).  From the above analysis we also
expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength
is independent of the specific form of the potential.

@@

To test the scaling of the equilibration time $t$ with respect to $H$
and $\delta$, we simulate the vesicle adhesion dynamics in a quiescent
flow.  Starting with two vesicles at a distance of twice the radius at
$t=0$, the long-range attraction pulls the vesicles together.
Figure~\ref{fig:qflow00}(a) plots the scaled adhesion energy versus time
with $H=0.1$ and three values of $\delta$ as labeled.  Adhesion energy
reaches minimum at equilibrium, and the scaled $E_{adh}$ evolves towards
zero at equilibrium.  The arrow indicate the times when the scaled
adhesion energy reaches within $1\%$ of equilibrium: $t\sim 96$ for
$\delta=0.2$, $t\sim 45$ for $\delta = 0.3$, and $t\sim 30$ for
$\delta=0.4$:
\begin{align*}
\frac{t_{\delta=0.2}}{t_{\delta=0.4}} \sim \frac{96}{30}=3.2 &\Longleftrightarrow \left(\frac{0.4}{0.2}\right)^{5/3}\sim 3.17,\\
\frac{t_{\delta=0.2}}{t_{\delta=0.3}} \sim \frac{96}{45}=2.1 &\Longleftrightarrow \left(\frac{0.3}{0.2}\right)^{5/3}\sim 2.
\end{align*}

The scaling with respect to adhesion strength $H$ is illustrated in
Figure~\ref{fig:qflow00}(b), where $\delta = 0.4$ and $H$ varies from
$0.1$ to $0.5$ as labeled.  Again the arrows indicate the times when the
equilibrium is reached within $1\%$:
\begin{align*}
\frac{t_{H=0.1}}{t_{H=0.5}} \sim \frac{34}{20} = 1.71 &\Longleftrightarrow \left(\frac{0.5}{0.1}\right)^{1/3}\sim 1.71.
\end{align*}
From the above results we conclude that the scaling in
equation~\eqref{eq:teq_scaling} captures the adhesion dynamics of two
vesicles that interact with each other via the adhesion force in
equation~\eqref{eq:Fh}.
